---
title: "01_Study_scope"
output: html_notebook
---

01_Study_scope

Code produces plots for Figure 3 (n.b plots further edited in Adobe Illustrator):
- Diseases studied 
- Genera studied (by disease)
- Type of study (field or experimental)
- Other parameters recorded (environmental)
- Other methods used to study coral disease 
- Study locations 

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(tidyr)
library(MetBrewer)
library(stringr)
library(circlize)
library(ggalluvial)

data <- read.csv("Data_collection_cleaned_cleaned.csv")

data_2 <- read.csv("V3_data_collection_cleaned.csv")

str(data)

reorder_size <- function(x) {
        factor(x, levels = names(sort(table(x), decreasing = TRUE)))
}

```
Disease type 

```{r}
disease <- data %>% group_by(Disease_type) %>% count(Disease_type)

# Seperate Disease_type by , and then put it into long format

disease_type <- data %>% select(Title,Publication_year, Disease_type) %>%
  separate(sep = ",", col = Disease_type, into = c("Disease_type_1","Disease_type_2")) %>%
  gather(key,value,3:4) %>%
  na.omit() %>% 
  group_by(value) 

# Proportions bar charts
ggplot(disease_type, aes(x = reorder_size(`value`))) +
        geom_bar(aes(y = (..count..)/sum(..count..)), fill = "pink1", colour = "pink2", alpha = 0.8) +
        xlab("Disease name") +
        scale_y_continuous(labels = scales::percent, name = "Proportion", breaks = seq(0,0.6,0.1)) +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_label(stat = 'count', aes(label = ..count..), position = position_fill(vjust =0.5), size = 2.2)


```
Disease genera combinations 

```{r}
species <- data %>% group_by(Taxa_count) %>% count(Taxa_count)

ggplot(species, aes(Taxa_count,n)) +
  geom_col() + 
  theme_bw() 

# Pull out species studied 

test <- data %>% 
  select(Title,Publication_year,Disease_genera_1) %>%
  separate(sep = ";", col = Disease_genera_1, into = c("x1","x2","x3","x4","x5","x6","x7","x8","x9","x10","x11","x12",
                                                     "x13","x14","x15","x16","x17","x18","x19","x20","x21","x22","x23","x24","x25","26")) %>%
  gather(key,value,3:28) %>%
  group_by(value) %>%
  separate(sep = ",", col = value, into = c("y1","y2"))
           
test2 <- test %>% 
  na.omit() #%>% 
  #filter(!row_number() %in% c(36:58,60:70,73:75,77,79:82))

test3 <- test2 %>% 
  mutate(across(where(is.character), str_trim))

# Full plot
  
 ggplot(test3, aes(x = reorder_size(y2))) +
        geom_bar(aes(y = (..count..)/sum(..count..)), fill = "pink1", colour = "pink2", alpha = 0.8) +
        xlab("Genera") +
        scale_y_continuous(labels = scales::percent, name = "Proportion", breaks = seq(0,0.6,0.1)) +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_label(stat = 'count', aes(label = ..count..), position = position_fill(vjust =0.5), size = 2.2)
 
 
 # Full plot
  
 ggplot(test3, aes(x = reorder_size(y1), fill = y2)) +
        geom_bar() +
        xlab("Genera") +
   theme_bw() +
     theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
   
        
```


Study type 

```{r}
studytype <- data %>% group_by(Study_type) %>% count(Study_type)

ggplot(studytype, aes(Study_type,n)) +
  geom_col() + 
  theme_bw()

View(study_type_data)

# Donut plot here

study <- data %>% group_by(Study_type) %>% count(Study_type)

# Compute percentages
study$fraction <- study$n / sum(study$n)

# Compute the cumulative percentages (top of each rectangle)
study$ymax <- cumsum(study$fraction)

# Compute the bottom of each rectangle
study$ymin <- c(0, head(study$ymax, n=-1))

# Compute label position
study$labelPosition <- (study$ymax + study$ymin) / 2

# Compute a good label
study$label <- paste0(study$Study_type, "\n value: ", study$count)

# Make the plot
ggplot(study, aes(ymax=ymax,ymin=ymin,xmax=4,xmin=3,fill=Study_type)) +
  geom_rect() +
  geom_text(x=2, aes(y=labelPosition, label=Study_type, color=Study_type, size=6)) + 
  scale_fill_manual(values=c("orchid4","plum3","orchid2")) +
  scale_color_manual(values=c("orchid4","plum3","orchid2")) +
  coord_polar(theta="y") +
  xlim(c(-1, 4)) +
  theme_void() 


```


Other environmental parameters recorded in field studies 

```{r}

# Specific to field 

studytype_field <- data %>% filter(Study_type == "Field") 

studytype_field <- data %>% filter(Study_type == "Field") %>%
  select(Title,Publication_year,Study_type, Environmental_parameters) %>%
  separate(sep = ",", col = Environmental_parameters, into = c("Env1","Env2")) %>%
  gather(key,value,4:5) %>%
  group_by(value) %>%
  na.omit()

ggplot(studytype_field, aes(x = reorder_size(value))) +
        geom_bar(aes(y = (..count..)/sum(..count..)), fill = "plum3", colour = "darkmagenta", alpha = 0.8) +
        xlab("Environmental parameters") +
        scale_y_continuous(labels = scales::percent, name = "Proportion", breaks = seq(0,0.7,0.1)) +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_label(stat = 'count', aes(label = ..count..), position = position_fill(vjust =0.6), size = 2.2)
  
```
Other methods used by field and experimental studies

```{r}

other_data <- data %>%
  select(Title,Publication_year,Other_methods) %>%
  separate(sep = ";", col = Other_methods, into = c("x1","x2")) %>%
  gather(key,value,3:4) %>%
  group_by(value) %>%
  na.omit()
  
# By study type 

other_data <- data %>% filter(Study_type == "Field")  %>%
  select(Title,Publication_year,Other_methods) %>%
  separate(sep = ";", col = Other_methods, into = c("x1","x2")) %>%
  gather(key,value,3:4) %>%
  group_by(value) %>%
  na.omit()

other_data_1 <- other_data %>% 
  mutate(across(where(is.character), str_trim))

ggplot(other_data_1, aes(x = reorder_size(value))) +
        geom_bar(aes(y = (..count..)/sum(..count..)), fill = "plum3", colour = "darkmagenta", alpha = 0.8) +
        xlab("Other data") +
        scale_y_continuous(labels = scales::percent, name = "Proportion", breaks = seq(0,0.6,0.1)) +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_label(stat = 'count', aes(label = ..count..), position = position_fill(vjust =0.5), size = 2.2)


# By study type 

other_data_2 <- data %>% filter(Study_type == "Experimental (lab)")  %>%
  select(Title,Publication_year,Other_methods) %>%
  separate(sep = ";", col = Other_methods, into = c("x1","x2")) %>%
  gather(key,value,3:4) %>%
  group_by(value) %>%
  na.omit()

other_data_3 <- other_data_2 %>% 
  mutate(across(where(is.character), str_trim))

ggplot(other_data_3, aes(x = reorder_size(value))) +
        geom_bar(aes(y = (..count..)/sum(..count..)), fill = "orchid4", colour = "darkorchid4", alpha = 0.8) +
        xlab("Other data") +
        scale_y_continuous(labels = scales::percent, name = "Proportion", breaks = seq(0,0.6,0.1)) +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_label(stat = 'count', aes(label = ..count..), position = position_fill(vjust =0.5), size = 2.2)


```


```{r}

species <- data %>% group_by(Taxa_count) %>% count(Taxa_count)

ggplot(species, aes(Taxa_count,n)) +
  geom_col() + 
  theme_bw() 

# Pull out species studied 

test <- data %>% 
  select(Title,Publication_year,Disease_genera_1) %>%
  separate(sep = ";", col = Disease_genera_1, into = c("x1","x2","x3","x4","x5","x6","x7","x8","x9","x10","x11","x12",
                                                     "x13","x14","x15","x16","x17","x18","x19","x20","x21","x22","x23","x24","x25","26")) %>%
  gather(key,value,3:28) %>%
  group_by(value) %>%
  separate(sep = ",", col = value, into = c("y1","y2"))
           
test2 <- test %>% 
  na.omit() #%>% 
  #filter(!row_number() %in% c(36:58,60:70,73:75,77,79:82))

test3 <- test2 %>% 
  mutate(across(where(is.character), str_trim))

# Full plot
  
 ggplot(test3, aes(x = reorder_size(y2))) +
        geom_bar(aes(y = (..count..)/sum(..count..)), fill = "pink1", colour = "pink2", alpha = 0.8) +
        xlab("Genera") +
        scale_y_continuous(labels = scales::percent, name = "Proportion", breaks = seq(0,0.6,0.1)) +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_label(stat = 'count', aes(label = ..count..), position = position_fill(vjust =0.5), size = 2.2)
 
 
 # Full plot
  
 ggplot(test3, aes(x = reorder_size(y1), fill = y2)) +
        geom_bar() +
        xlab("Genera") +
   theme_bw() +
     theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
   
        
  
```

Location of sampling 

```{r}

data$Study_location_reef_area


 ggplot(data, aes(x = reorder_size(Study_location_reef_area))) +
        geom_bar(aes(y = (..count..)/sum(..count..)), fill = "pink1", colour = "pink2", alpha = 0.8) +
        xlab("Genera") +
        scale_y_continuous(labels = scales::percent, name = "Proportion", breaks = seq(0,0.6,0.1)) +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_label(stat = 'count', aes(label = ..count..), position = position_fill(vjust =0.5), size = 2.2) +
   coord_flip()

```


