---
title: "R Notebook"
output: html_notebook
---

02_Coral_histopathology_methodologies

Code produces plots for Figure 4 (n.b plots further edited in Adobe Illustrator):
- Study aim 
- Sample type 
- Data type 
- Examination categories for histopath (split into other associated orgs, host qualitative/quantitative)
- Disease by other associated org plot (sankey)
- Disease by other host terminology plot (sankey)

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(tidyr)
library(MetBrewer)
library(stringr)
library(circlize)
library(ggalluvial)

data <- read.csv("Data_collection_cleaned_cleaned.csv")

data_2 <- read.csv("V3_data_collection_cleaned.csv")

str(data)

reorder_size <- function(x) {
        factor(x, levels = names(sort(table(x), decreasing = TRUE)))
}

```


Study aim for histology (made into venn diagram in illustrator)

```{r}

ggplot(data, aes(x = reorder_size(Focus))) +
        geom_bar(aes(y = (..count..)/sum(..count..)), fill = "pink1", colour = "pink2", alpha = 0.8) +
        xlab("Aim") +
        scale_y_continuous(labels = scales::percent, name = "Proportion", breaks = seq(0,0.8,0.1)) +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_label(stat = 'count', aes(label = ..count..), position = position_fill(vjust =0.7), size = 2.2) + 
  coord_flip()
          
```

Assessment of sample type  (made into venn diagram in illustrator)

```{r}

sample_type <- data %>% 
  select(Title,Publication_year,Sample_types) %>%
  separate(sep = ",", col = Sample_types, into = c("x1","x2","x3","x4","x5")) %>%
  gather(key,value,3:7) %>%
  na.omit() %>% 
  mutate(across(where(is.character), str_trim))

# Full plot
  
 ggplot(sample_type, aes(x = reorder_size(value))) +
        geom_bar(aes(y = (..count..)/sum(..count..)), fill = "pink1", colour = "pink2", alpha = 0.8) +
        xlab("Genera") +
        scale_y_continuous(labels = scales::percent, name = "Proportion", breaks = seq(0,0.8,0.1)) +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_label(stat = 'count', aes(label = ..count..), position = position_fill(vjust =0.7), size = 2.2)+
   coord_flip()
          
 
ggplot(data, aes(x = reorder_size(Sample_types))) +
        geom_bar(aes(y = (..count..)/sum(..count..)), fill = "pink1", colour = "pink2", alpha = 0.8) +
        xlab("Genera") +
        scale_y_continuous(labels = scales::percent, name = "Proportion", breaks = seq(0,0.8,0.1)) +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_label(stat = 'count', aes(label = ..count..), position = position_fill(vjust =0.7), size = 2.2)+
   coord_flip()

```

Data type (made into venn diagram in illustrator)

```{r}

data_type <- data %>% 
  select(Title,Publication_year,Data_type_2) %>%
  separate(sep = ",", col = Data_type_2, into = c("x1","x2")) %>%
  gather(key,value,3:4) %>%
  na.omit() %>% 
  mutate(across(where(is.character), str_trim))

# Full plot
  
 ggplot(data_type, aes(x = reorder_size(value))) +
        geom_bar(aes(y = (..count..)/sum(..count..)), fill = "pink1", colour = "pink2", alpha = 0.8) +
        xlab("Genera") +
        scale_y_continuous(labels = scales::percent, name = "Proportion", breaks = seq(0,0.8,0.1)) +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_label(stat = 'count', aes(label = ..count..), position = position_fill(vjust =0.7), size = 2.2)+
   coord_flip()
 
 # Full plot
  
 ggplot(data, aes(x = reorder_size(Data_type_2))) +
        geom_bar(aes(y = (..count..)/sum(..count..)), fill = "pink1", colour = "pink2", alpha = 0.8) +
        xlab("Genera") +
        scale_y_continuous(labels = scales::percent, name = "Proportion", breaks = seq(0,0.8,0.1)) +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_label(stat = 'count', aes(label = ..count..), position = position_fill(vjust =0.7), size = 2.2)+
   coord_flip()

```

What was examined?

```{r}

examined<- data %>% 
  select(Title,Publication_year,Examined) %>%
  separate(sep = ",", col = Examined, into = c("x1","x2","x3","x4","x5","x6","x7","x8","x9","x10","x11","x12","x13","x14","x15","x16","x17","x18","x19","x20","x21","x22","x23","x24","x25","x26")) %>%
  gather(key,value,3:28) %>%
  group_by(value) %>% 
  na.omit() %>% 
  mutate(across(where(is.character), str_trim))

# Filter out those with "associated" 

examined_associated <- examined %>% filter(grepl("Associated",value)) %>% 
  mutate(across(where(is.character), str_trim))

# Full plot
  
 ggplot(examined, aes(x = reorder_size(value))) +
        geom_bar(aes(y = (..count..)/sum(..count..)), fill = "pink1", colour = "pink2", alpha = 0.8) +
        xlab("Genera") +
        scale_y_continuous(labels = scales::percent, name = "Proportion", breaks = seq(0,0.4,0.1)) +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_label(stat = 'count', aes(label = ..count..), position = position_fill(vjust =0.3), size = 2.2)

# Full plot
  
 ggplot(examined_associated, aes(x = reorder_size(value))) +
        geom_bar(aes(y = (..count..)/sum(..count..)), fill = "pink1", colour = "pink2", alpha = 0.8) +
        xlab("Genera") +
        scale_y_continuous(labels = scales::percent, name = "Proportion", breaks = seq(0,0.5,0.1)) +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_label(stat = 'count', aes(label = ..count..), position = position_fill(vjust =0.4), size = 2.2)
   
```


```{r}

examined_host <- examined %>% filter(!grepl("Associated",value)) %>% 
  mutate(across(where(is.character),str_trim)) %>% 
  na.omit()

# Full plot
  
 ggplot(examined_host, aes(x = reorder_size(value))) +
        geom_bar(aes(y = (..count..)/sum(..count..)), fill = "pink1", colour = "pink2", alpha = 0.8) +
        xlab("Genera") +
        scale_y_continuous(labels = scales::percent, name = "Proportion", breaks = seq(0,0.5,0.1)) +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_label(stat = 'count', aes(label = ..count..), position = position_fill(vjust =0.4), size = 2.2)

```

Quantitative assessments 

```{r}

data$Quantiative
quantitative <- data %>% 
  select(Title,Publication_year,Quantiative) %>%
  separate(sep = ",", col = Quantiative, into = c("x1","x2","x3")) %>%
  gather(key,value,3:5) %>%
  group_by(value) %>% 
  na.omit() %>% 
  mutate(across(where(is.character), str_trim))

# Filter out those with no value

# Full plot
  
 ggplot(quantitative, aes(x = reorder_size(value))) +
        geom_bar(aes(y = (..count..)/sum(..count..)), fill = "pink1", colour = "pink2", alpha = 0.8) +
        xlab("Genera") +
        scale_y_continuous(labels = scales::percent, name = "Proportion", breaks = seq(0,0.3,0.1)) +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_label(stat = 'count', aes(label = ..count..), position = position_fill(vjust =0.3), size = 2.2)

```

# Seperate out diseases 

```{r}
disease <- data %>% group_by(Disease_type) %>% count(Disease_type)

# Seperate Disease_type by host examination , and then put it into long format

disease_type <- data %>% select(Title,Publication_year,Examined, Disease_type) %>%
  separate(sep = ",", col = Disease_type, into = c("Disease_type_1","Disease_type_2")) %>%
  gather(key,value,4:5) %>%
  na.omit() %>% 
  group_by(value) 

```


```{r}

# Seperate Disease_type by other organisms examined, and then put it into long format

disease_type <- data %>% select(Title,Publication_year,Examined, Disease_type) %>%
  separate(sep = ",", col = Disease_type, into = c("Disease_type_1","Disease_type_2")) %>%
  gather(key,value,4:5) %>%
  na.omit() %>% 
  group_by(value) 

colnames(disease_type)[colnames(disease_type) == "value"] ="disease_type"

examined_disease <- disease_type %>% 
  dplyr::select(Title,Publication_year,disease_type,Examined) %>%
  separate(sep = ",", col = Examined, into = c("x1","x2","x3","x4","x5","x6","x7","x8","x9","x10","x11","x12","x13","x14","x15","x16","x17","x18","x19","x20","x21","x22","x23","x24","x25","x26","x27")) %>%
  gather(key,value,4:30) %>%
  group_by(value) %>% 
  na.omit() %>% 
  mutate(across(where(is.character), str_trim))


# Filter out associated organisms (host only)

examined_host_disease <- examined_disease %>% filter(!grepl("Associated",value)) %>% 
  mutate(across(where(is.character),str_trim)) %>% 
  na.omit()

# Filter out associated organisms 

examined_associated_disease <- examined_disease %>% filter(grepl("Associated",value)) %>% 
  mutate(across(where(is.character), str_trim))

```

Sankey diagrams!

Host condition/reproduction
```{r}
# Calculate frequency 

examined_host_disease_freq <-  examined_host_disease%>%
  group_by(disease_type,value) %>%
  count() %>%
  na.omit()

# Make a colour list for diseases 

colour <- c()

ggplot(examined_host_disease_freq, aes(y = n,axis1 = disease_type, axis2 = value)) +
  geom_alluvium() +
  geom_stratum() +
 geom_label(stat = "stratum", aes(label = after_stat(stratum))) +
  scale_fill_manual(values = colors) +
  #scale_x_continuous(breaks = 1:3, labels = c("Hair", "Eye", "Sex")) +
  #coord_flip() +
  theme_void()



```

Other associated organisms 

```{r}
# Calculate frequency 

examined_other_disease_freq <- examined_associated_disease %>%
  group_by(disease_type,value) %>%
  count() %>%
  na.omit()

# Make a colour list for diseases 

colour <- c()

ggplot(examined_other_disease_freq, aes(y = n,axis1 = disease_type, axis2 = value)) +
  geom_alluvium() +
  geom_stratum() +
 geom_label(stat = "stratum", aes(label = after_stat(stratum))) +
  scale_fill_manual(values = colors) +
  #scale_x_continuous(breaks = 1:3, labels = c("Hair", "Eye", "Sex")) +
  #coord_flip() +
  theme_void()

```


